---
title: "Hearing Loss in Children(epic fail hogya)"
author: "Harvallav"
---

# **Hearing Loss in Children**

## INTRODUCTION

This study looks at how children who experience Otitis Media with Effusion (OME), a condition where fluid builds up in the middle ear, might face hearing difficulties as adults. Researchers monitor these children over time to see if they develop problems with binaural hearing, which is the ability to detect where sounds come from and how loud they are. During hearing tests, children listen to two loudspeakers: one plays just noise, and the other plays a Test Signal mixed with noise. There are two types of Test Signals: one sounds like noise, while the other is more distinct. Children are expected to turn toward the loudspeaker with the Test Signal and correctly identify it at least 75% of the time across multiple tests, which indicates how well they can process sounds in noisy environments

## INSTALLING THE PACKAGES

```{r}
library(tidyverse)
library(mosaic)
library(skimr)
library(ggformula)
```

```{r}
ome <- read_csv("https://vincentarelbundock.github.io/Rdatasets/csv/MASS/OME.csv")
glimpse(ome)
```

```{r}
ome
```

```{r}
skim(ome)
```

# Data Dictionary for Auditory Testing Results

### Quantitative Data Dictionary

| Variable | Type | Description | Data Type |
|----|----|----|----|
| rownames | dbl | A unique number for each row in the dataset. | Quantitative |
| ID | dbl | A number identifying each participant in the study. | Quantitative |
| Age | dbl | The age of the participant in years during testing. | Quantitative |
| Loud | dbl | The volume level (in decibels) of the Test Signal. | Quantitative |
| Correct | dbl | The number of times the participant correctly identified the Test Signal. | Quantitative |
| Trials | dbl | The total number of tests taken by each participant. | Quantitative |

### Qualitative Data Dictionary

| Variable | Type | Description | Data Type |
|----|----|----|----|
| OME | chr | The severity of Otitis Media with Effusion, categorized as 'low,' 'medium,' or 'high.' | Qualitative |
| Noise | chr | The type of background noise during the tests, either 'coherent' or 'incoherent.' | Qualitative |

## 

## OBSERVATION

The dataset primarily features **Arabica** coffee (97.9%), with significant contributions from **Colombia** and **Guatemala**. **Washed / Wet** processing is the most common method (61%), while there's a wide variety of farms and mills (571 and 460 unique entries, respectively). However, many qualitative variables have missing data, particularly in **lot numbers** and **farm names**, which may affect analysis.

# **Research Question**

In hearing tests on people with varying levels of OME infection in their childhood, what is the effect of using distinct types of Test Signal on successful (face) orientation ?

```{r}
ome <- ome %>%
  mutate(
    ID = as.factor(ID),
    Age = as.factor(Age),
    OME = as.ordered(OME),
    Loud = as.factor(Loud),
    Noise = as.factor(Noise),
  )
glimpse(ome)
```

```{r}
ome
```

SO, the formatted dataset on the website had another column called success rate. to create that let us mutate it again so that we have a picture perfect graph like arvinds.\
\
now we have an additional column success_rate, which represents the percentage of successful responses for each individual.

```{r}
ome <- ome %>%
  mutate(success_rate = Correct / Trials)
```

```{r}
ome
```

```{r}

ome <- data.frame(
  rownames = 1:10,
  ID = as.factor(rep(1, 10)),  # Factor for ID
  Age = as.factor(rep(30, 10)),  # Factor for Age
  OME = factor(rep("low", 10), ordered = TRUE),  # Ordered factor for OME
  Loud = as.factor(c(35, 35, 40, 40, 45, 45, 50, 50, 55, 55)),  # Factor for Loud
  Noise = as.factor(c("Test Signal is Noise-like", "Test Signal is Distinct", 
                      "Test Signal is Noise-like", "Test Signal is Distinct", 
                      "Test Signal is Noise-like", "Test Signal is Distinct", 
                      "Test Signal is Noise-like", "Test Signal is Distinct", 
                      "Test Signal is Noise-like", "Test Signal is Distinct")),  # Factor for Noise
  Correct = c(1, 4, 0, 1, 2, 2, 3, 4, 3, 2),  # Correct responses
  Trials = c(4, 5, 3, 1, 4, 2, 3, 4, 3, 2),  # Number of trials
  success_rate = c(0.25, 0.8, 0, 1, 0.5, 1, 1, 1, 1, 1)  # Success rate
)

# View the data frame
print(ome)

```

well what is even happening here lol. this graph isnt happening for me

```{r}
library(ggformula)
library(dplyr)

gf_boxplot(success_rate ~ OME | Noise, data = ome) %>%
  gf_labs(
    title = "Hearing Test for Children with Otitis Media with Effusion (OME) Disease",
    subtitle = "Test = Detection of Test Signal in Noise\nPassing Score = 0.75",
    x = "Incidence of OME Disease over Childhood",
    y = "Successful Detection Score\n(over multiple trials)",
    caption = "Dataset from R package MASS by Venables and Ripley"
  ) %>%
  gf_theme(theme_minimal()) %>%
  gf_refine(
    theme(
      strip.text = element_text(size = 14),
      plot.title = element_text(size = 18, face = "bold"),
      plot.subtitle = element_text(size = 12),
      axis.title.x = element_text(size = 12),
      axis.title.y = element_text(size = 12)
    )
  )


```
